---
title: "Statistical Analyses for Brigette Haram Lake Thurmond data"
output:
#  pdf_document:
#    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Introduction
========================================================
Data set, statistical analyses, graphics and inference concerning bird abundance, species richness in the presence/absence of Hydrilla.

Citation:
R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing,
  Vienna, Austria. URL https://www.R-project.org/.

Distribution
========================================================
This page is not yet public on rpubs:
http://rpubs.com/puruckertom/lake_thurmond

The git repo holding this code is available at:
https://github.com/puruckertom/bharam_lakerussell

Computational environment
========================================================

```{r eval=FALSE, echo=FALSE}
#library(installr) #windows only currently
#updateR()
```
Version and installed libraries.
```{r eval=TRUE, echo=TRUE}
verbose_output = TRUE

if(verbose_output){
  print(Sys.info()[4])
  R.Version()$version.string
}
#check to see if needed packages are installed
list.of.packages <- c("ggplot2", "dplyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
#load relevant packages
library(MASS)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(BSDA)

if(verbose_output){
  print("list of loaded packages: ")
  print((.packages()))
}
```

```{r eval=TRUE, echo=FALSE}
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  lake_russell <- "k:\\git\\bharam_lakerussell\\"
}

if(Sys.info()[4]=="stp-air.local"){
  lake_russell <- path.expand("~/git/bharam_lakerussell/")
}

lr_data <- paste(lake_russell, "data_in/", sep="")
```

Observational data
========================================================
Original file with observational data: 
https://github.com/puruckertom/bharam_lakerussell/blob/master/data_in/lakerussell_data.txt

If only x is given, or if both x and y are given and paired is TRUE, a Wilcoxon signed rank test of the null that the distribution of x (in the one sample case) or of x - y (in the paired two sample case) is symmetric about mu is performed.

```{r eval=TRUE, echo=TRUE}
##################
#the data sets
##################
#everything
file.exists(paste(lr_data,"lakerussell_data.csv",sep=""))
lr_obs <- read.table(paste(lr_data,"lakerussell_data.csv",sep=""), header = TRUE, sep = ",")
str(lr_obs)
```

```{r eval=TRUE, echo=TRUE}
lr.nohydrilla <- lr_obs[which(lr_obs$Hydrilla==1),]
lr.hydrilla <- lr_obs[which(lr_obs$Hydrilla==2),]
```

Statistics
========================================================
wilcox test as a generalization of the binomial test, bootstrapped test statistic by randomly sampling 
the hydrilla to pair with the 22 nonhydrilla samples for each realization, repeated 1000 times
to create a distribution of the test statistic

the null is that the median of the differences of bird abundance and/or density (between paired 
hydrilla and nonhydrilla samples) is equal to zero, ties are therefore equal to zero and
should not be discarded

wilcoxon rank rum (= mann-whitney u) can also handle unequal sample sizes

the sign.test does not reliably return a htest object as it is supposed to:
http://r.789695.n4.nabble.com/SIGN-test-td4654434.html

Inference
========================================================
```{r eval=TRUE, echo=TRUE}
pvalues.birddensity <- rep(NA, 1000)
for(i in 1:1000){
  nohydrilla.birddensity <- lr.nohydrilla$Bird.ha
  hydrilla.birddensity <- sample(lr.hydrilla$Bird.ha,size=22)
  #exact suppresses error warnings
  x <- wilcox.test(nohydrilla.birddensity, hydrilla.birddensity,paired=TRUE,exact=FALSE)$p.value
  y <- SIGN.test(nohydrilla.birddensity, hydrilla.birddensity, md = 0, alternative = "two.sided", conf.level = 0.95)
  pvalues.birddensity[i] <- x
}
hist(pvalues.birddensity)
paste("median = ", median(pvalues.birddensity))
```
```{r eval=TRUE, echo=TRUE}
pvalues.birdabundance <- rep(NA, 1000)
for(i in 1:1000){
  nohydrilla.birdabundance <- lr.nohydrilla$Species.ha
  hydrilla.birdabundance <- sample(lr.hydrilla$Species.ha,size=22)
  delta.birdabundance <- hydrilla.birdabundance - nohydrilla.birdabundance
  #exact suppresses error warnings
  #z <- wilcox.test(nohydrilla.birdabundance, hydrilla.birdabundance,paired=TRUE,exact=FALSE)$p.value
  #y <- SIGN.test(nohydrilla.birdabundance, hydrilla.birdabundance, md = 0, alternative = "two.sided", conf.level = 0.95)
  x <- binom.test(18,22,p=0.5,alternative="two.sided",conf.level=0.95)
  pvalues.birdabundance[i] <- x
}
hist(pvalues.birdabundance)
paste("median = ", median(pvalues.birdabundance))
```